Traffic aware Resource Allocation schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} systems

### 2014-10-01 Report/wsrm_formulation.tex
temporary_version
double_ready

In this approach, we present an algorithm to solve \eqref{eqn-3} for the transmit precoders in a centralized way by using the idea of alternating optimization and successive convex approximation. Using \eqref{eq:SINR}, we can reformulate the problem defined in \eqref{eqn-3} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}  |^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  \tilde{N}_0 + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad &\eqref{eqn-4.3} \IEEEyessubnumber
\end{IEEEeqnarray}
}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
	\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \vspace{-0.25cm}\\
	\text{subject to} &\quad& \gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \eqspace \label{eqn-6.2} \\
	&\quad& \beta_{l,k,n} \geq \tilde{N}_0 + \sum_{\mathclap{(j,i) \neq (l,k)}} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-6.3} \\
	&\quad& \text{and } \eqref{eqn-4.3} \IEEEyessubnumber
\end{IEEEeqnarray}
}
where \me{\tilde{\mbf{u}}_{l,k,n} \triangleq \{\mvec{w}{l,k,n}^\herm, \mvec{H}{b_k,k,n}, \mvec{m}{l,k,n},\beta_{l,k,n}\}} be the vector which needs to be identified for the optimal allocation and \me{\tilde{N}_0 = N_0\|\mvec{w}{l,k,n}\|^2} be the effective noise variance. In this formulation, we relaxed the equality constraint in \eqref{eq:SINR} by the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3}. However, this step without loss of optimality leads to the same solution, since the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3} are active for an optimal solution, following the same arguments as those in \cite{tran2012fast}. Intuitively, \eqref{eqn-6.2} denotes the \ac{SINR} constraint for \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} gives an upper bound for the total interference seen by the user \me{k \in \mathcal{U}_b}, denoted by the variable \me{\beta_{l,k,n}}. Similar to the \ac{WSRM} problem in \cite{tran2012fast}, the problem can be shown to be NP-hard even for the single antenna case. The reformulation in \eqref{eqn-6} allows a tractable solution as presented below. First, we note that the constraints \eqref{eqn-4.3} are convex with involved variables. Thus, we only need to deal with \eqref{eqn-6.2} and \eqref{eqn-6.3}. Towards this end, we resort to the traditional coordinate descent technique by fixing the linear receivers, and finding the optimal transmit beamformers. Recall the original coordinate descent method assumes that the optimization variables belong to disjoint sets (Cartesian product of sets, to be precise) \cite{xu2013block}.

@@

The proposed algorithm is referred as \acl{QM} \ac{JSFRA} scheme with a per \ac{BS} power constraint which is outlined in Algorithm \ref{algo-1}. The iterative procedure repeats until the improvement on the objective is less than a predetermined tolerance parameter or the maximum number of iterations is reached. Instead of initializing \me{\tilde{\mbf{u}}_{l,k,n}} arbitrarily to a feasible point, transmit precoders can also be initialized with any feasible point \me{\tilde{\mbf{m}}_{l,k,n}}, which is then used to find \me{\tilde{\mbf{u}}_{l,k,n}} in an efficient manner as briefed in Algorithm \ref{algo-1}. For a fixed receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{SCA} iteration is carried out until convergence or for the predefined iterations, say \me{J_{\max}} for the optimal transmit precoders \me{\mvec{m}{l,k,n}}. Next, the receive beamformers are updated based on either \eqref{opt-rx} or \eqref{eqn-10} using the fixed transmit precoders \me{\mvec{m}{l,k,n}}. This procedure is carried out until convergence of the queue deviation or for fixed number of iterations defined by \me{I_{\max}} as outlined in Algorithm \ref{algo-1}.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0} and transmit precoders \me{\tilde{\mbf{M}}_{k,n}} randomly satisfying the total power constraint \eqref{eqn-4.3}}
 update \me{\mvec{W}{k,n}} and \me{\tilde{\mbf{u}}_{l,k,n}} using \eqref{eqn-10} and \eqref{eqn-8} using \me{\tilde{\mbf{M}}_{k,n}}\;
 \Repeat{Queue convergence or \me{i \geq I_{\max}}}{
 initialize \me{j = 0}\;
 \Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
 solve for the transmit precoders \me{\mvec{m}{l,k,n}} using \eqref{eqn-9}\;
 update the constraint set \eqref{eqn-8} with \me{\tilde{\mbf{u}}_{l,k,n}} and \me{\mvec{m}{l,k,n}} using \eqref{eqn-wsrm-expr}\;
 $j = j + 1$\;
 }
 update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-9--1} or \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
 $i = i + 1$\;
 }
 \caption{Algorithm of \acs{JSFRA} scheme}
 \label{algo-1}
\end{algorithm}

\subsubsection*{Convergence}

>>>>

In this approach, we present an algorithm to solve \eqref{eqn-3} to obtain the transmit precoders in a centralized manner by using the idea of alternating optimization and successive convex approximation. Using \eqref{eq:SINR}, we can reformulate the problem defined in \eqref{eqn-3} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\ \beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}  |^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  \tilde{N}_0 + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad &\eqref{eqn-4.3} \IEEEyessubnumber
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
	\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \vspace{-0.25cm}\\
	\text{subject to} &\quad& \gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \eqspace \label{eqn-6.2} \\
	&\quad& \beta_{l,k,n} \geq \tilde{N}_0 + \sum_{\mathclap{(j,i) \neq (l,k)}} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-6.3} \\
	&\quad& \text{and } \eqref{eqn-4.3} \IEEEyessubnumber
\end{IEEEeqnarray}
}
where \me{\tilde{\mbf{u}}_{l,k,n} \triangleq \{\mvec{w}{l,k,n}^\herm, \mvec{H}{b_k,k,n}, \mvec{m}{l,k,n},\beta_{l,k,n}\}} be the vector which needs to be identified for the optimal allocation and \me{\tilde{N}_0 = N_0\|\mvec{w}{l,k,n}\|^2} be the effective noise variance. In this formulation, we relaxed the equality constraint in \eqref{eq:SINR} by the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3}. However, this step without loss of optimality leads to the same solution, since the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3} are active for an optimal solution, following the same arguments as those in \cite{tran2012fast}. Intuitively, \eqref{eqn-6.2} denotes the \ac{SINR} constraint for \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} gives an upper bound for the total interference seen by the user \me{k \in \mathcal{U}_b}, denoted by the variable \me{\beta_{l,k,n}}. Similar to the \ac{WSRM} problem in \cite{tran2012fast}, the problem can be shown to be NP-hard even for the single antenna case. The reformulation in \eqref{eqn-6} allows a tractable solution as presented below. First, we note that the constraints \eqref{eqn-4.3} are convex with involved variables. Thus, we only need to deal with \eqref{eqn-6.2} and \eqref{eqn-6.3}. Towards this end, we resort to the traditional coordinate descent technique by fixing the linear receivers, and finding the optimal transmit beamformers. Recall the original coordinate descent method assumes that the optimization variables belong to disjoint sets (Cartesian product of sets, to be precise) \cite{xu2013block}.

@@

The proposed algorithm is referred as \acl{QM} \ac{JSFRA} scheme with a per \ac{BS} power constraint which is outlined in Algorithm \ref{algo-1}. The iterative procedure repeats until the improvement on the objective is less than a predetermined tolerance parameter or the maximum number of iterations is reached. Instead of initializing \me{\tilde{\mbf{u}}_{l,k,n}} arbitrarily to a feasible point, transmit precoders can also be initialized with any feasible point \me{\tilde{\mbf{m}}_{l,k,n}}, which is then used to find \me{\tilde{\mbf{u}}_{l,k,n}} in an efficient manner as briefed in Algorithm \ref{algo-1}. For a fixed receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{SCA} iteration is carried out until convergence or for the predefined iterations, say \me{J_{\max}} for the optimal transmit precoders \me{\mvec{m}{l,k,n}}. Next, the receive beamformers are updated based on either \eqref{opt-rx} or \eqref{eqn-10} using the fixed transmit precoders \me{\mvec{m}{l,k,n}}. This procedure is carried out until convergence of the queue deviation or for fixed number of iterations defined by \me{I_{\max}} as outlined in Algorithm \ref{algo-1}.
{\allowdisplaybreaks
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0} and transmit precoders \me{\tilde{\mbf{M}}_{k,n}} randomly satisfying the total power constraint \eqref{eqn-4.3}}
 update \me{\mvec{W}{k,n}} and \me{\tilde{\mbf{u}}_{l,k,n}} using \eqref{eqn-10} and \eqref{eqn-8} using \me{\tilde{\mbf{M}}_{k,n}}\;
 \Repeat{Queue convergence or \me{i \geq I_{\max}}}{
 initialize \me{j = 0}\;
 \Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
 solve for the transmit precoders \me{\mvec{m}{l,k,n}} using \eqref{eqn-9}\;
 update the constraint set \eqref{eqn-8} with \me{\tilde{\mbf{u}}_{l,k,n}} and \me{\mvec{m}{l,k,n}} using \eqref{eqn-wsrm-expr}\;
 $j = j + 1$\;
 }
 update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-9--1} or \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
 $i = i + 1$\;
 }
 \caption{Algorithm of \acs{JSFRA} scheme}
 \label{algo-1}
\end{algorithm}
}
\subsubsection*{Convergence}
